The impact of gas bulk rotation on the Lyman-$\alpha$ line

### 2014-01-15 paper/ms.tex
two more items in the conclusions

We present results of radiative transfer calculations to measure the
impact of gas bulk rotation on the morphology of the Lyman $\alpha$
emission line in distant galaxies. We model a galaxy as a sphere with
a homogeneous mixture of dust and hydrogen at a constant
temperature. These spheres have a solid-body rotation with maximum
velocities in the range $0-300$ \kms and neutral hydrogen optical
depths in the range $\tau_{\rm H}=10^{5}-10^{7}$. We also consider two
kinds of spatial distribution for the radiation sources with respect
to the sphere: central and homogeneous. Our main finding is that the
line width and the intensity at the line's center increase with
rotational velocity. For homogeneously distributed sources the line
becomes single peaked at large rotational velocities. Under the same
conditions the escape fraction increases $\sim 30\%$. For radiation
sources located off-center, the line morphology presents a range of
single, double and triple peaks. We show how these results are useful
to interpret recent spectroscopic results of distant $z\sim 2-3$ star
forming galaxies. 
\end{abstract}

@@

This paper illustrates for the first time the main effects of rotation
in the morphology of the \ly emission line in star forming galaxies. 
The results are based on the study of an homogeneous sphere
with solid body rotation. We explored a range of models by varying
the rotational speed, hydrogen optical depth, dust optical depth and
initial distribution of \ly photons with respect to the gas
density. As a cross-validation, we obtain our results from two
independently developed Monte-Carlo radiative transfer codes.

Our main result is that rotation clearly impact the \ly line
morphology. Double peaked lines can make transitions into single
peaked lines as the rotational velocity increases. Additional
off-centered emission result into triple peaked lines for certain 
values of optical depth and velocities.

\begin{itemize}
\item For central sources we find that the number of scatterings
does not significantly decrease with rotation, this is due to the 
fact that the vast majority of scatterings events are resonant, 
during which the mean free path is very short and the effect of gas
bulk rotation is not enough to affect \ly photons.
The escape fraction does not depend on the rotational velocity neither
the position of the peaks, but there is an increase 
in the flux on the central parts of the line for high rotational 
velocities.

@@

%We have quantified the results in different aspects.

%These changes in the number of scatterings have a direct bearing on
%the escape fraction in the presence of dust.  In Section
%\ref{sec:escapefraction} we find that the escape fraction increases in
%$20\%-30\%$ for the homogeneous source distribution. In the centrally
%located emission, the rotational effects are negligible on the escape
%fraction.

%We also studied the effect of rotation in the position of the peaks in
%the line. We find that in the regions with central emission the peaks
%do not shift as rotation increases. However, in the homogeneous models
%the double peak merges into a single peak.

@@

JNGC thanks the support of Universidad de los Andes given as teaching assistant, 
Also to the  ISAAC 2012 summer school held University of California San Diego which provide computer time
 where some of this models where ran.

>>>>

We present results of radiative transfer calculations to measure the
impact of gas bulk rotation on the morphology of the Lyman $\alpha$
emission line in distant galaxies. We model a galaxy as a sphere with
a homogeneous mixture of dust and hydrogen at a constant
temperature. These spheres have a solid-body rotation with maximum
velocities in the range $0-300$ \kms and neutral hydrogen optical
depths in the range $\tau_{\rm H}=10^{5}-10^{7}$. We also consider two
kinds of spatial distribution for the radiation sources with respect
to the sphere: central and homogeneous. Our main finding is that the
line width and the intensity at the line's center increase with
rotational velocity. For homogeneously distributed sources the line
becomes single peaked at rotational velocities larger than the line
width in the static case. Under the same
conditions the escape fraction increases $\sim 30\%$. For radiation
sources located off-center, the line morphology presents a range of
single, double and triple peaks. We show how these results are useful
to interpret recent spectroscopic results of distant $z\sim 2-3$ star
forming galaxies.

\end{abstract}

@@

This paper quantifies for the first time in the literature the effects
of rotation in the morphology of the \ly emission line in star forming
galaxies.  The results are based on the study of an homogeneous sphere
of gas with solid body rotation. We explore a range of models by varying
the rotational speed, hydrogen optical depth, dust optical depth and
initial distribution of \ly photons with respect to the gas
density. As a cross-validation, we obtain our results from two
independently developed Monte-Carlo radiative transfer codes.

Our main result is that rotation clearly impacts the \ly line
morphology. Double peaked lines can make transitions into single
peaked lines as the rotational velocity increases. In the case of
off-centered emission result into triple peaked lines for certain
values of optical depth and rotational velocities.

Quantitatively the main results of our study are summarized in the
following items.

\begin{itemize}

\item The line width increases with rotational velocity. The line with
  approximately follows the functional form $W_{\rm 1/2}^2 = W_{\rm
    1/2,0}^2 + (V_{\rm max}/\lambda)^2$ where $V_{\rm 1/2,0}$ indicates the line
  width for the static case. $\lambda$ is a constat determined from
  the radiative transfer results it is $\lambda_{\rm c}=XX$ and
  $\lambda_{\rm h}=YY$ for the central and homogeneous source
  distribution, respectively.

\item A single peaked line emerges for high rotational velocities in
  the case of homgeneously distributed sources. The cases of single
  peaked lines occur when the rotational velocity is larger than the
  line width measured in the static case.

\item For central sources we find that the number of scatterings
does not significantly decrease with rotation, this is due to the 
fact that the vast majority of scatterings events are resonant, 
during which the mean free path is very short and the effect of gas
bulk rotation is not enough to affect \ly photons.
The escape fraction does not depend on the rotational velocity neither
the position of the peaks, but there is an increase 
in the flux on the central parts of the line for high rotational 
velocities.

@@

@@

JNGC acknowledges financial support from Universidad de los
Andes.

JEFR acknowledges financial support from Vicerrectoria de
Investigaciones at Universidad de los Andes through a FAPA grant.

We thank the International Summer School on AstroComputing
2012 organized by the University of California High-Performance
AstroComputing Center (UC-HiPACC) for providing computational
resources where some of the calculations were done.
